\section{Conclusions and Future Work}
The paper presents the VR system which is an HJ runtime system for
verification. The runtime supports \texttt{async}, \texttt{finish},
\texttt{isolated}, and future data-driven forms of \texttt{async} in
the HJ language and is small enough to manual inspect for
correctness. It runs standalone as an alternative HJ runtime system
and in the JPF model checker without any modification to the runtime or JPF. When run in the JPF model checker, JPF
is able to prove an HJ program data-race free. Such a proof ensures HJ
safety guarantees such as deadlock freedom, deterministic execution,
and the program is able to serialize. The paper further compares the
VR system to HJ runtime systems. It also shows the performance gains
in JPF between the default scheduler and the optimized scheduler.

Future work is to extend the VR system to support the \texttt{forall}, \texttt{forasync}, data-driven programming with \texttt{await}, \texttt{next}, \texttt{phaser} constructs, and exceptions which are
also part of the HJ semantics. There is also a need to formalize the
behavior of the new optimized default scheduler factory in JPF and
prove that is sound and complete for proving data-race freedom in HJ
programs. The scheduler will need to be further extended for the new
HJ constructs, though the default schedule would still be sound and
complete for data-race freedom. Finally, future work needs to address
scalability in large HJ programs as the number of schedules to
consider becomes rather prohibitive in these programs.
